---
title: "individual_final"
author: "jiaying Ning"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
#loading package
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(ggridges)
library(usdata)
library(usmap)
library(ggplot2)
library(patchwork)

library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
    fig.asp = .6,
  out.width = "90%"

)

theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete =scale_fill_viridis_d
```



#### introducing first dataset "covid_impact_on_airport_traffic"
```{r}
#import data for covid_impact_on_airport_traffic
#Main Problem: join data, have date and state match for both data
# Transportation and Covid have different way to representing date(2020-01-01 in transportation date and 01/01/2020 in Covid data)
transportation = 
   read_csv("./data/covid_impact_on_airport_traffic.csv" ) %>%
  janitor::clean_names() %>%
  filter(country != "Australia" & country != "Chile" & country != "Canada") %>% #only leave United States data 
   separate(date, c("year","month","day"), sep = "([-])") %>%  # I re-arrange the date information so that it matched the date format in Covid data
    mutate(date = paste(month,day,year,sep = "/")) %>% # I re-arrange the date information so that it matched the date format in Covid data
  relocate(date) %>% 
  select(-year,-month,-day,-version,-aggregation_method,-state,-country) %>% #delete variable that is not in our interest
    rename(state = iso_3166_2) %>% #rename state variable so that we can combine two data 
  mutate(state=gsub("US-","",state)) # reformat state variable, delete prefix of US-



```

#### introducing first dataset "United_States_Covid-19_Cases_and_Deaths_by_State_over_Time"
```{r}
Covid =  
  read_csv("./data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv" ) %>%
  janitor::clean_names() %>%
  filter(consent_cases == "Agree" | consent_deaths == "Agree")%>% #need to decided whether this step is necessary
  select("submission_date":"pnew_death") %>% #select variable of interest, need to look further into which variable we are interested in and way
    rename(date = submission_date)  # rename date variable so that we can match data accordingly
  
```

#### Joining two dataset
```{r}

Covid_transport_data = 
  left_join(transportation, Covid, by = c("date")) %>% #left join two data, by date
  filter(state.y == state.x) #filter the data so that we only leave the data that have matching date and state
  
```


#### Introducing a third dataset
```{r}

Mobility = 
   read_csv("./data/2020_US_Region_Mobility_Report.csv" ) %>%
  janitor::clean_names() %>%
   separate(date, c("year","month","day"), sep = "([-])") %>%
  mutate(
    state = state2abbr(sub_region_1),
  date = paste(month,day,year,sep = "/"))  %>%
   select(-metro_area,-country_region_code,-day,-month,-year) %>%
  filter(!is.na(sub_region_1))
  
```


```{r}
Mobility_clean = 
Mobility %>%
  mutate(date = fct_inorder(date),
         date = as.Date(date, "%m/%d/%y")) %>%
  select(-iso_3166_2_code,-country_region,-census_fips_code,-sub_region_2,-sub_region_1) %>%
  relocate(date,state) 
```

```{r}
Covid_clean =
Covid %>%
  mutate(date = fct_inorder(date),
         date = as.Date(date, "%m/%d/%y")) %>%
  select(-conf_cases,-prob_cases,-pnew_case,-conf_death,-prob_death,-pnew_death)
  
```




```{r}
Covid_mobility_data=
  left_join( Covid_clean,Mobility_clean, by = c("date")) %>% #left join two data, by date
  filter(state.x==state.y) %>%
  mutate(state=state.x) %>%
  select(-state.x,-state.y)
```


```{r}
rm(Covid)
rm(Mobility)
rm(Mobility_clean)
rm(Covid_clean)


```


```{r}

#There seems to be multiple date for each date and state, try to take mean of it 
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) 
  
```


```{r}

#TRY to make plot about the relationship between mean_transit station percent change and date
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_transit,color = state)) +
 geom_line(alpha=0.3) +
 scale_x_date(date_breaks = "1 month")
  
```

```{r}
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = sum_newcases,color = state)) +
 geom_point(alpha=0.3) +
 scale_x_date(date_breaks = "1 month")
  
```


```{r}
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline) & new_case >0) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = sum_newcases,y = mean_transit,color = state)) +
 geom_point(alpha=0.3)

```


```{r}
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline) & new_case >0) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_tot_cases = sum(tot_cases),
            n = n()) %>%
  ggplot(aes(x = mean_transit,y = sum_tot_cases,color = state)) +
 geom_point(alpha=0.3)

```

```{r}
#use plotly to compare the mean transit among each state
Covid_mobility_data %>%
  group_by(state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline) & new_case >0) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            n = n()) %>%
  mutate(state = fct_reorder(state, mean_transit)) %>% 
   plot_ly(x = ~state, y = ~mean_transit, color = ~state, type = "bar", colors = "viridis")
```

```{r}
#use plotly to compare the mean transit among each state
Covid_mobility_data %>%
  group_by(state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline) & new_case >0) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_new_cases = sum(new_case),
            n = n()) %>%
  mutate(state = fct_reorder(state, sum_new_cases)) %>% 
   plot_ly(x = ~state, y = ~sum_new_cases, color = ~state, type = "bar", colors = "viridis")
```


```{r}
#use plotly to compare the mean transit among each state
Covid_mobility_data %>%
  group_by(state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline) & new_case >0) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_new_cases = mean(new_case),
            n = n()) %>%
  mutate(state = fct_reorder(state, mean_new_cases)) %>% 
   plot_ly(x = ~state, y = ~mean_new_cases, color = ~state, type = "bar", colors = "viridis")
```
```

```{r}
plot_ILCA_MEAN_NEWCASE=
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)& state == c("LA","IL")) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_newcases,color = state)) +
 geom_point(alpha=0.3) +
  geom_smooth()+
 scale_x_date(date_breaks = "1 month") +
    facet_wrap(~state)

```

```{r}
plot_ILCA_MEAN_Transit=
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)& state == c("LA","IL")) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_transit,color = state)) +
 geom_point(alpha=0.3) +
  geom_smooth()+
 scale_x_date(date_breaks = "1 month") +
    facet_wrap(~state)
```


```{r}
plot_ILCA_MEAN_Transit/plot_ILCA_MEAN_NEWCASE
```


```{r}
Covid_mobility_data %>%
  group_by(date,state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)& state == c("LA","IL")) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_totalcases = sum(tot_cases),
            n = n()) %>%
  ggplot(aes(x = date,y = sum_totalcases,color = state)) +
 geom_point(alpha=0.3) +
  geom_smooth()+
 scale_x_date(date_breaks = "1 month") +
    facet_wrap(~state)
```

```{r}

#comparing only by date
Covid_mobility_data %>%
  group_by(date) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = sum_newcases,color="light pink")) +
 geom_point(aes(size=mean_transit,alpha=0.3))+
 scale_x_date(date_breaks = "1 month")+   
  labs(titles = "number of newcases daily",
        y = "New Cases"  )
  
```
```{r}
#comparing only by date
#TRANSIT
Covid_mobility_data %>%
  group_by(date) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
             sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_transit,color="light pink")) +
 geom_point(aes(size=sum_newcases,alpha=0.3))+
 scale_x_date(date_breaks = "1 month")+
    labs(titles = "change in transit after covid",
        y = "change in transit from baseline"  )
```

```{r}
Covid_mobility_data %>%
  group_by(date) %>%
   filter(!is.na(retail_and_recreation_percent_change_from_baseline)) %>%
  summarise(mean_retail_recreation = mean(retail_and_recreation_percent_change_from_baseline),
             sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_retail_recreation,color="light pink")) +
 geom_point(aes(size=sum_newcases,alpha=0.3))+
 scale_x_date(date_breaks = "1 month")+  
  labs(titles = "change in retial and recreation after covid",
        y = "change in retail and recreation from baseline"  )
  
```


```{r}

Covid_mobility_data %>%
  group_by(date) %>%
   filter(!is.na(residential_percent_change_from_baseline)) %>%
  summarise(mean_residential = mean(residential_percent_change_from_baseline),
             sum_newcases = sum(new_case),
            n = n()) %>%
  ggplot(aes(x = date,y = mean_residential,color="light pink")) +
 geom_point(aes(size=sum_newcases,alpha=0.3))+
 scale_x_date(date_breaks = "1 month")+  
  labs(titles = "change in retial and recreation after covid",
        y = "change in retail and recreation from baseline"  )
  
```

```{r}
#comparing only by state
Covid_mobility_data %>%
  group_by(state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) 

```


```{r}
mean_Covid_mobility_data =
  Covid_mobility_data %>%
  group_by(state) %>%
   filter(!is.na(transit_stations_percent_change_from_baseline)) %>%
  summarise(mean_transit = mean(transit_stations_percent_change_from_baseline),
            mean_newcases = mean(new_case),
            n = n()) 
```




